Solar wind predictions for the Parker Solar Probe orbit

### 2017-08-16 paperMVVB/sections/content.tex
text flow streamlining

%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1962 \citep{Neugebauer1966} \hl{+ ref. to Russian measurements}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in the case of Voyager~1 as far away as \SI{138}{\au} in July 2017\footnote{\url{https://voyager.jpl.nasa.gov/}}, having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Until today various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with Wind \hl{(ref.)}, ACE \hl{(ref.)} and DSCOVR \hl{(ref.)} still orbiting around the L1 point \SI{1.5}{million \km} ahead of Earth in the sunward direction. Additional measurements at other distances were provided by planetary missions to Venus and Mercury, such as PVO \hl{(ref.)} or MESSENGER \hl{(ref.)}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind at polar latitudes \citep{McComas1998}. The nearest in situ solar wind measurements to date were obtained by the Helios mission. The in 1974 launched Helios~1 spacecraft reached distances of \SI{0.31}{\au}, Helios~2 launched two years later and approached the Sun up to \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). {\color{Red2} In comparison, the ESA Solar Orbiter mission with a planned launch in 2019 will have a closest perihelion of \SI{0.289}{\au} \hl{(ref.)}.}

@@

In order to help optimize the WISPR and PSP preplanning of the science operations knowledge of the expected solar wind environment is needed. For this purpose the solar wind environment is extrapolated down to the closest perihelion of \SI{9.86}{\Rs} distance to the Sun using in situ solar wind data from the Helios probes and near \SI{1}{\au} data from various satellites compiled in the NASA/GSFC OMNI solar wind database.

%definition of sw environment
%section{Solar wind}
Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, {\color{red} such as ...intermittent reconnection at the top of streamers and pseudostreamers, coronal hole boundaries, ... (\hl{ref.}).} The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of coronal mass ejections (CMEs), the faster ones driving shock waves ahead \hl{(ref.)}. Their rate follows the activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

%lognormal fitting
\section{Frequency distributions of the solar wind parameters}
\label{sec:frequency_distribution}

The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on the phase of the solar activity cycle. Hence, for deriving characteristic frequency distributions for the solar wind parameters, measurements over long-term time spans are needed. The abundance of the near-Earth hourly OMNI data set is ideally suited for this purpose, because it spans to date almost five solar cycles.

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind plasma and magnetic field data collected by various satellites since 1963. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
It should be noted that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016, did not show significant differences in our results.

%data binning
According to the OMNI data precision and maximal parameter ranges we specify the bin sizes to \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 ICME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

	\caption{Resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions from near \SI{1}{\au} OMNI hourly data. For the velocity also the fit parameters of the double lognormal function (\ref{eq:double_lognormal_fit_function}) are listed, as well as the median and mean values of the resulting velocity fit. The mean absolute errors and sums of absolute residuals are also listed. The values in brackets are the estimated standard deviations of each fit parameter.}

@@

		S[table-format = 1.2e+2]

@@

		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Balance}	&\multicolumn{1}{c}{\multirow{2}{*}{MAE}}	&\multicolumn{1}{c}{SAR}\\
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{$x_\text{med}$}	&\multicolumn{1}{c}{$x_\text{avg}$}	&\multicolumn{1}{c}{$c$}	&\multicolumn{1}{c}{}	&\multicolumn{1}{c}{[\%]}\\

@@

		\multicolumn{2}{l}{Magnetic field}	&5.661(16)	&6.164(18)	&--	&5.51e-4	&6.83\\
		\multicolumn{2}{l}{Velocity}	&4.085(19)	&4.183(20)	&--	&1.80e-3	&18.69\\
		\multicolumn{2}{l}{Density}	&5.276(24)	&6.484(34)	&--	&5.49e-4	&6.48\\
		\multicolumn{2}{l}{Temperature}	&7.470(17)	&11.301(32)	&--	&8.71e-5	&5.78\\

@@

			&\multicolumn{1}{c}{$W_\text{II}$}	&4.16(14)\tablefootmark{b}	&4.42(14)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&3.98e-4	&4.20\\

@@

From visual inspection, the resulting curves describe the shape of the magnetic field strength, density and temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. However, for the velocity the fit function appears not to be as good in describing the measured distribution’s more complex shape around its maximum and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit listed in Table~\ref{tab:lognormal_fit_parameters}, being almost three times larger than those from the other parameters.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fit's median and mean values are indicated as well. The insets show zoomed-in frequency axes.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
This more complex fit function is more accurately in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{The velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.
%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well, but differ for the extreme values, mainly caused by CME events.
The simple lognormal fit functions underestimate the frequency of the solar wind parameters in their high value tails, except for the temperature’s tail which is overestimated as seen by the insets in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.

\section{Solar activity dependence of the solar wind frequency distributions}
\label{sec:solar_activity_variations}
In the next step we investigate how the long-term solar wind distribution functions presented in the previous section depend on general solar activity. Therefore we examine their correlation with the sunspot number, being a commonly used long-term solar activity index and determine the time lags with the highest correlation coefficients.

@@

Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} shows yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number. The reason for correlating the SSN to the solar wind median values is, because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance. The correlation coefficients of the solar wind parameters with the yearly SSN shown in the bottom part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} are calculated for time lags back to \num{-15}~years to cover a time span longer than a solar cycle.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{The solar wind parameter yearly medians derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

@@

The highest correlation coefficient of 0.728 to the SSN is found for the magnetic field strength which has no time lag. This finding is anticipated because the SSN is found to be directly proportional to the evolution of the photospheric magnetic flux \citep{Smith2003}.
Velocity and temperature show a lag time of 3~years with peak correlation coefficients of 0.453 and 0.540. The density with a correlation coefficient of 0.468 has a lag time of 6~years, which is in agreement with the by \citet{Bougeret1984} reported density anticorrelation to the SSN.	%p.~406\\
% maximal correlation coefficients and their lag times:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\
As expected, the correlation coefficients' amplitudes of all parameters decline with increasing lag time and show a frequency of about 11~years.

>>>>

%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1962 \citep{Neugebauer1966} \hl{+ ref. to Russian measurements}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}} (July 2017), having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Until today various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with Wind \hl{(ref.)}, ACE \hl{(ref.)} and DSCOVR \hl{(ref.)} still orbiting around the L1 point \SI{1.5}{million \km} ahead of Earth in the sunward direction. Additional measurements at other solar distances were provided by planetary missions to Venus and Mercury, such as PVO \hl{(ref.)} or MESSENGER \hl{(ref.)}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind even at polar latitudes \citep{McComas1998}. The Sun-nearest in situ solar wind measurements to date were obtained by the Helios mission. The in 1974 launched Helios~1 spacecraft reached distances of \SI{0.31}{\au}, Helios~2 launched two years later and approached the Sun up to \SI{0.29}{\au} \citep{Rosenbauer1977}.
%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rs), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). {\color{Red2} In comparison, the ESA Solar Orbiter mission with a planned launch in 2019 will have a closest perihelion of \SI{0.289}{\au} \hl{(ref.)}.}

@@

In order to help optimize the WISPR and PSP preplanning of the science operations knowledge of the expected solar wind environment is needed. For this purpose the solar wind environment is extrapolated down to the closest perihelion of \SI{9.86}{\Rs} distance to the Sun using in situ solar wind data from the Helios probes and near \SI{1}{\au} data from various satellites compiled in the OMNI solar wind database.

%definition of sw environment
%section{Solar wind}
Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966,Schwenn1983}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses' out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types \citep{Schwenn1983}, is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, {\color{red} such as ...intermittent reconnection at the top of streamers and pseudostreamers, coronal hole boundaries, ... (\hl{ref.}).} The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.
Embedded in the slow and fast solar wind streams are transient flows of coronal mass ejections (CMEs), the faster ones driving shock waves ahead \hl{(ref.)}. Their rate follows the solar activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

@@

%lognormal fitting
\section{Frequency distributions of the solar wind parameters}
\label{sec:frequency_distribution}
The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on the phase of the solar activity cycle. Hence, for deriving characteristic frequency distributions for the solar wind parameters, measurements over long-term time spans are needed. The abundance of the near-Earth hourly OMNI data set is ideally suited for this purpose, because it spans to date almost five solar cycles.

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind magnetic field and plasma data collected by various satellites since 1963. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the different parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
It should be noted that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016 did not show significant differences in our results.

%data binning
According to the OMNI data precision and maximal parameter ranges we specify bin sizes of \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature. The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}.
The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean data values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).
Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 ICME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.

@@

	\caption{Resulting fit coefficients from the fitting of the lognormal function (\ref{eq:single_lognormal_fit_function}) to the shape of the solar wind parameters' frequency distributions from near \SI{1}{\au} OMNI hourly data. For the velocity also the fit parameters of the double lognormal function (\ref{eq:double_lognormal_fit_function}) are listed, as well as the median and mean values of the resulting velocity fit. The mean absolute errors and sums of absolute residuals are also listed. The values in brackets are the estimated standard deviations of the fit parameters.}

@@

		S[table-format = 2.3]

@@

		\multicolumn{2}{l}{\multirow{2}{*}{Parameter}}	&\multicolumn{1}{c}{Median\tablefootmark{a}}	&\multicolumn{1}{c}{Mean\tablefootmark{a}}	&\multicolumn{1}{c}{Balance}	&\multicolumn{1}{c}{\multirow{1}{*}{MAE}}	&\multicolumn{1}{c}{SAR}\\
		\multicolumn{2}{l}{}	&\multicolumn{1}{c}{$x_\text{med}$}	&\multicolumn{1}{c}{$x_\text{avg}$}	&\multicolumn{1}{c}{$c$}	&\multicolumn{1}{c}{[$10^{-4}$]}	&\multicolumn{1}{c}{[\%]}\\

@@

		\multicolumn{2}{l}{Magnetic field}	&5.661(16)	&6.164(18)	&--	&5.51	&6.83\\
		\multicolumn{2}{l}{Velocity}	&4.085(19)	&4.183(20)	&--	&18.0	&18.69\\
		\multicolumn{2}{l}{Density}	&5.276(24)	&6.484(34)	&--	&5.49	&6.48\\
		\multicolumn{2}{l}{Temperature}	&7.470(17)	&11.301(32)	&--	&0.871	&5.78\\

@@

			&\multicolumn{1}{c}{$W_\text{II}$}	&4.16(14)\tablefootmark{b}	&4.42(14)\tablefootmark{b}	&\multicolumn{1}{c}{--}	&3.98	&4.20\\

@@

From visual inspection, the resulting fit curves describe the shape of the magnetic field strength, density and temperature distributions well, as can be seen in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. However, for the velocity the fit function appears not to be as good in describing the measured distribution’s more complex shape around its peak and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit listed in Table~\ref{tab:lognormal_fit_parameters}, being almost three times larger than those from the other parameters.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits derived from the hourly OMNI data set. The histograms have bins of \SI{0.5}{\nT}, \SI{10}{\km\per\s}, \SI{1}{\per\cm\cubed} and \SI{10000}{\K}. The fits' median and mean values are indicated as well. The insets show zoomed-in frequency axes.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution, we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.
The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
This more complex fit function is more accurate in describing the velocity's frequency distribution as shown in Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{The velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset has a zoomed-in frequency axis.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
Thus in the following sections we keep the double lognormal ansatz for all velocity frequency fits.

%transition
For the bulk of the solar wind these static lognormal functions describe the parameters' distributions well, but differ for the extreme values, mainly caused by CME events.
The simple lognormal fit functions underestimate the frequency of the solar wind parameters in their high value tails, except for the temperature’s tail which is overestimated as seen in the insets of Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. The velocity's compositional lognormal fit only slightly overestimates its tail as seen in the inset of Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.
The slow and fast part contribute almost equally ($c \approx 0.5$) to the long-term velocity distribution function.

{\color{red} discuss high value zoom figures; read in Veselovsky2010}

\section{Solar activity dependence of the solar wind frequency distributions}
\label{sec:solar_activity_variations}
In the next step we investigate how the long-term solar wind distribution functions presented in the previous section depend on general solar activity. Therefore we examine their correlation with the sunspot number, being a commonly used long-term solar activity index, and determine the time lags with the highest correlation coefficients.

@@

Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} shows yearly medians of the solar wind parameters and the yearly SSN together with the solar cycle number. The reason for correlating the SSN to the solar wind median values is because the position of a lognormal function is defined by its median. The data are averaged to yearly values to avoid seasonal effects during the Earth’s orbit around the Sun caused by its variations in solar latitude and distance.
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/OMNI_yearly_ssn_correlation_c_plot.pdf}}
	\caption{The solar wind parameter yearly medians derived from OMNI data and the yearly SSN from the \citet{sidc} with solar cycle number (top). Their correlation coefficients with the yearly SSN are calculated for time lags back to -15 years (bottom).}
	\label{fig:OMNI_yearly_ssn_correlation_c_plot}
\end{figure}

@@

The correlation coefficients of the solar wind parameters with the yearly SSN shown in the bottom part of Fig.~\ref{fig:OMNI_yearly_ssn_correlation_c_plot} are calculated for time lags back to \num{-15}~years to cover a time span longer than a solar cycle. As expected, the correlation coefficients' amplitudes of all parameters decline with increasing time lag and show a frequency of about 11~years. The highest correlation coefficient of 0.728 to the SSN is found for the magnetic field strength, it has no time lag. This finding is anticipated because the SSN is found to be directly proportional to the evolution of the photospheric magnetic flux \citep{Smith2003}.
Velocity and temperature show a lag time of 3~years with peak correlation coefficients of 0.453 and 0.540. The density with a correlation coefficient of 0.468 has a time lag of 6~years, which is in agreement with the by \citet{Bougeret1984} reported density anticorrelation to the SSN.	%p.~406\\
% maximal correlation coefficients and their lag times:\\
% magnetic field strength: lag 0, 0.728\\
% velocity: lag -3, 0.453\\
% density: lag -6, 0.468\\
% temperature: lag -3, 0.540\\
